Thesis Title

### 2022-02-01 chapters/accronims.tex
Progress on introduction

Global Burden of Disease (GBD)
World Health Organization (WHO)
Global Priority Pathogens (GPP)
antimicrobial resistance (AMR)

>>>>

Global Burden of Disease (GBD)
World Health Organization (WHO)
Global Priority Pathogens (GPP)
antimicrobial resistance (AMR)
T2 Magnetic Resonance (T2MR)
polymerase chain reaction (PCR)
pulsed field gel electrophoresis (PFGE)
Multilocus sequence typing (MLST)
basepairs (bp)
Real-time quantitative PCR (qPCR)
Lateral flow immunoassays (LFIA)
ribossomal RNA (rRNA)

### 2022-02-01 chapters/introduction.tex
Progress on introduction

Clinical microbiology is a discipline focused on rapidly characterising pathogen samples in order to direct the management of individual infected patients (diagnostic microbiology) and monitor the epidemiology of infectious disease (public health microbiology), including the detection of outbreaks and infection prevention. Globally, ... FINANTIAL BURDEN!

@@

In the early 2000s, Multilocus sequence typing (MLST) was proposed as a portable, universal, and definitive method for characterising bacteria \citep{maiden_multilocus_2006}. Instead of enzyme restriction of bacteria DNA, separation of the restricted DNA bands using a pulsed-field electrophoresis chamber, followed by clonal assignment of bacteria based on PFGE banding patterns, MLST relies on the amplification through PCR sequences of internal fragments of house-keeping genes (usually 5 to 7), approximately 450-500 basepairs (bp) in size, followed by its the sequence, usually my Sanger methods. For each house-keeping gene, the different sequences present within a bacterial species are assigned as distinct alleles and, for each isolate, the alleles at each of the (usually) seven loci define the allelic profile or sequence type \citep{larsen_multilocus_2012}. As with PFGE, different schemes, defining what house-keeping gene fragnments are used, are available depending of the species. Unlike PFGE, the provision of freely accessible, curated databases of MLST nucleotide sequence data enables the direct comparison of bacterial isolates, providing the basis of a common language for bacterial typing \citep{maiden_multilocus_2006}. MLST schemes for over 100 organisms having been published and made freely available (\url{https://pubmlst.org/organisms}, \cite{jolley_open-access_2018})

@@

The traditional approaches to laboratory diagnosis of viral infections have been (1) direct detection in patient material of virions, viral antigens, or viral nucleic acids, (2) isolation of virus in cultured cells, followed by identification of the isolate, and (3) detection and measurement of antibodies in the patient’s serum (serology). \citep{burrell_laboratory_2017}

@@

@@

Currently, the detection and diagnosis of parasite infections rely on several laboratory methods in addition to clinical symptoms, clinical history, travel history, and geographic location of patient. serology-based assays, MOLECULAR-BASED ASSAYS	, PROTEOMICS, microscopy  \citep{ndao_diagnosis_2009}
    %\item The plethora of new antiviral agents and a more sophisticated understanding of how these should be used;
    %\item An increased awareness of the clinical value of, and demand for, prompt information about viral loads, viral sequence data, and antiviral resistance information.

>>>>

Clinical microbiology is a discipline focused on rapidly characterising pathogen samples in order to direct the management of individual infected patients (diagnostic microbiology) and monitor the epidemiology of infectious disease (public health microbiology), including the detection of outbreaks and infection prevention. According to WHO's Global Expenditure on Health report from 2000 to 2019, of the 51 countries that reported health spending by disease and condition, an average of 37\% of health spending went to infectious and parasitic diseases, corresponding to the largest share of health spending \citep{world_health_organization_global_2021}. About 21\% of total health spending went to three major infectious diseases — HIV/AIDS (9\%), tuberculosis (1\%) and malaria (11\%) — and 16\% went to other infectious and parasitic diseases. On average, 70\% of external aid for health went to infectious and parasitic diseases in the 51 low and middle income countries. Of the \$54.8 billion estimated disbursed for health in 2020, \$13.7 billion (25\%) was targeted toward the COVID-19 health response \citep{micah_tracking_2021}.

@@

In the early 2000s, Multilocus sequence typing (MLST) was proposed as a portable, universal, and definitive method for characterising bacteria \citep{maiden_multilocus_2006}. Instead of enzyme restriction of bacteria DNA, separation of the restricted DNA bands using a pulsed-field electrophoresis chamber, followed by clonal assignment of bacteria based on PFGE banding patterns, MLST relies on the amplification through PCR sequences of internal fragments of house-keeping genes (usually 5 to 7), approximately 450-500 basepairs (bp) in size, followed by its the sequence, usually my Sanger methods. For each house-keeping gene, the different sequences present within a bacterial species are assigned as distinct alleles and, for each isolate, the alleles at each of the (usually) seven loci define the allelic profile or sequence type \citep{larsen_multilocus_2012}. As with PFGE, different schemes, defining what house-keeping gene fragments are used, are available depending of the species. Unlike PFGE, the provision of freely accessible, curated databases of MLST nucleotide sequence data enables the direct comparison of bacterial isolates, providing the basis of a common language for bacterial typing \citep{maiden_multilocus_2006}. MLST schemes for over 100 organisms having been published and made freely available (\url{https://pubmlst.org/organisms}, \cite{jolley_open-access_2018})

@@

The traditional approaches to laboratory diagnosis of viral infections have been (1) direct detection in patient material of virions, viral antigens, or viral nucleic acids, (2) isolation of virus in cultured cells, followed by identification of the isolate, and (3) detection and measurement of antibodies in the patient’s serum (serology) \citep{burrell_laboratory_2017}. Viral diagnostics is therefore generally organised into two primary categories, indirect and direct detection, depending on the method used.

Indirect detection methods involve the propagation of virus particles via their introduction to a suitable host cell line (virus isolation), as viruses rely on host organisms to replicate. This is a relatively slow diagnostic method, sometimes taking weeks for the virus to propagate, usually followed by miscroscopy for its identification, or more commonly, through molecular methods with an agent which detects a virus-associated protein, such as an antibody \citep{cassedy_virus_2021}.

Direct detection methods negate the need for virus propagation, detecting the virus directly from the suspect source through nucleic acid and immunological methods. PCR and reverse transcription-PCR are widely applied methods for the detection of both DNA and RNA viruses, respectively, driven by an increased awareness of the clinical value of, and demand for, prompt information about viral loads, viral sequence data, and potential antiviral resistance information. %CITATION
Real-time quantitative PCR (qPCR) remains the front line tool in aetiological diagnosis, measuring the production of the target amplicon throughout the reaction and providing quantitative results with high specificity and sensibility, albeit with a significant cost due to sophisticated apparatus despite high-throughput systems being widely established \citep{cassedy_virus_2021}. Immunoassays employ singular-epitope specificity antibodies as the primary means to detect viruses within a sample, and provides as a much more cost-efficient alternative to nucleic acid detection \citep{cassedy_virus_2021}. One major application is seroprevalence assays, an essential technique for identifying patients who have been exposed to a virus (historical exposure), detect asymptomatic infection or evaluate vaccine efficacy  \citep{chan_determining_2021, bobrovitz_global_2021}. Lateral flow immunoassays (LFIA) are extensively used for detecting virus-associated protein directly from source through labelled antibodies binding to their cognate antigens, usually read by way of a colour change at a test line. Besides being very cost effective, LFAIs have a turnaround time of minutes and the colour change can be observed with the naked eye, therefore facilitating rapid diagnosis but its results are limited to semi-quantitative and it does not typically achieve sensitivity comparable to nucleic-acid detection \citep{estrela_lateral_2016, cassedy_virus_2021, di_nardo_ten_2021}.

@@

Invasive fungal infections are a prevalent disease in pulmonary and critical care medicine, and they frequently pose diagnostic challenges, with laboratory diagnosis of fungal infections being identified as a key area in need of development of a clinical practise guideline \citep{haydour_diagnosis_2019}. Both serologic and molecular approaches are currently in use, with the first including not only immunodiffusion and complement fixation, both of which remain a cornerstone for fungal diagnostic testing, but also lateral-flow assays for fungal antigen detection. Molecular techniques for fungal identification both from culture and directly from patient specimens, include nucleic acid probes, mass spectrometry-based methods, and nucleic acid amplification testing \citep{ramanan_laboratory_2017}. Classic fungal culture continues to be relevant and necessary, with most detection and identification methods requiring growth in culture, from phenotypic identification and staining techniques, to nucleic acid hybridisation probes and MALDI-TOF mass spectrometry.

Molecular methods of fungal identification directly from the suspect source are particularly attractive for fungi that fail to produce characteristic morphologic characters in a timely fashion, but molecular databases for fungi tend to lag behind bacterial databases in terms of maturity \citep{ramanan_laboratory_2017}. qPCR testing  is one of the most frequently used
molecular methods in clinical microbiology, even days before the onset of positive culture result \citep{kourkoumpetis_polymerase_2012}. The ribossomal RNA (rRNA) gene cluster is frequently used as target for fungi identification, as well as other targets to evaluate antifungal resistance, guiding therapy, which lowers patient mortality and decrease unnecessary antifungal treatment, improving treatment-associated costs and avoiding toxicity \citep{kourkoumpetis_polymerase_2012}.

@@

Currently, the detection and diagnosis of parasite infections rely on several laboratory methods in addition to clinical symptoms, clinical history, travel history, and geographic location of patient. serology-based assays, molecular-based assays and microscopy  \citep{ndao_diagnosis_2009}. The use of the microscopy for the morphology-based diagnosis of disease remains the reference standard in parasitology although identification and matching of the life-cycle stages, despite its low cost, is very challenging even with expert knowledge \citep{blasco-costa_molecular_2016}. LFAIs has been used successfully, overcoming many of the disadvantages of conventional microscopy by providing simple, quick, cost-effective, and sensitive and accurate identification of parasites, particularly in low-income countries, but are available for a relatively low number of organisms, low sensitivity and reports of cross-reactivity \citep{momcilovic_rapid_2019, robert-gangneux_epidemiology_2012, noauthor_false_nodate}.

As observed in the diagnosis of bacterial, viral and fungal infections, molecular methods by nucleic acid amplification through qPCR has been increasingly adopted, particularly through the commercialisation of multiplex with high sensitivity and accuracy \citep{momcilovic_rapid_2019, wong_molecular_2014}. The key limitation of molecular detection is the technological expertise and expense which are usually lacking in the field setting at highly endemic areas.
